---
title: "Train Occupancy Prediction in Belgium"
author: "Hanyong Xu, Jia Yuan, Yibing Zheng"
date: "12/2019"
output:
  html_document:
    toc: TRUE
    toc_float: true
    code_folding: hide
    prettydoc::html_pretty:
    theme: cosmo
    highlight: tango
---
<style>
.superbigimage{
overflow-x:scroll;
white-space: nowrap;
}
.superbigimage img{
max-width: none;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	cache = FALSE,
	results = FALSE
)
options(scipen=99)
```

## 1. Introduction


```{r}
library(tidyverse)
library(sf)
library(QuantPsyc)
library(RSocrata)
library(viridis)
library(caret)
library(spatstat)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
library(nnet)
library(reshape2)
library(class)
library(MASS)
library(dplyr)
library(tidyr)
library(lubridate)
library(plotly)
library(ggmap)
library(data.table)
library(gganimate)
library(corrplot)
library(stargazer)
library(tigris)
library(ckanr)
```

```{r}
mapTheme <- function(base_size = 10) {
  theme(
    text = element_text(color = "black"),
    plot.title = element_text(size = 12,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    axis.title = element_blank(),
    strip.background = element_rect(fill = "white", color = "white"),
    strip.text.x = element_text(size = 8),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank()
  )
}

plotTheme <- theme(
  plot.title =element_text(size=12),
  plot.subtitle = element_text(size=8),
  plot.caption = element_text(size = 6),
  axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 10),
  # Set the entire chart region to blank
  panel.background=element_blank(),
  plot.background=element_blank(),
  axis.ticks=element_blank())
q5 <- function(variable) {as.factor(ntile(variable, 5))}
qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                          c(.01,.2,.4,.6,.8), na.rm=T)
  }
}
plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}
nn_function <- function(measureFrom,measureTo,k) {
  
  nn <-   
    FNN::get.knnx(measureTo, measureFrom, k)$nn.dist
  
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint)
  
  return(output)  
}
```

### Aggregate Data Analysis

Charts about stop-level ridership change before and after CapRemap
```{r}
data <- read.csv('C:/Users/HanyongXu/Documents/Me/grad/Spring_2020/MUSA801/Data/MUSA Data - Stop Ridership Aggregated.csv')

data = data %>% 
  rename(STOP_ID = ï..STOP_ID,)

table(data$YEAR_ID)
data.y18 <- data %>%
  subset(data$YEAR_ID == "2018")

max(data.y18$AVERAGE_ON)
```

```{r}
as.numeric(data$STOP_ID)

data.y18$MONTH_ID <- factor( data.y18$MONTH_ID, levels = c( "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12") )
data.y.1 <- data.y18 %>%
  subset(data.y18$STOP_ID <= 1050 & data.y18$AVERAGE_ON <= 30)

summary(data.y18$AVERAGE_ON)
```

##Maps about stops and ridership
Import School District shapefile
```{r}
TrusteeDist <- 
  st_read("C:/Users/HanyongXu/Documents/Me/grad/Spring_2020/MUSA801/Data/Trustee_Boundaries/Trustee.shp")%>%
  st_transform(2278) 

ggplot()+
  geom_sf(data = TrusteeDist)

```
Join stops and school districts
```{r}
data.sf <-
  data %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, agr = "constant") %>%
  st_transform(2278)

Austin.sf<-
  data.sf%>%
  st_join(dplyr::select(TrusteeDist,TRUSTEE))

data.y.Trustee18 <- Austin.sf %>%
  subset(Austin.sf$YEAR_ID == "2018")

```

```{r}
data.y.Trustee18$MONTH_ID <- factor(data.y.Trustee18$MONTH_ID, levels = c( "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12") )

plot.Trustees <- 
  as.data.frame(data.y.Trustee18) %>%
  group_by(TRUSTEE, MONTH_ID) %>% 
   summarize(BOARD_Count = sum(AVERAGE_ON)) %>%
    ggplot(aes(MONTH_ID,BOARD_Count)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~TRUSTEE, scales="free", ncol=3) +
      ylim(0,32000) +
      labs(title="Average Daily Ridership by Independent School District in Austin",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership in the District")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.Trustees
```

Making charts for Stops in TrusteeDistrict 1
```{r}
data.y.18.S1 <- data.y.Trustee18 %>%
  subset(data.y.Trustee18$TRUSTEE == "1")
summary(data.y.18.S1$AVERAGE_ON)


plot.S1.1 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID <= 1099) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,30) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.1

```
```{r}
plot.S1.2 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1099 & STOP_ID <= 1123) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,20) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.2
```
```{r}
plot.S1.3 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1123 & STOP_ID <= 1141) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.3
```
```{r}
plot.S1.4 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1141 & STOP_ID <= 1324) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,200) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.4
```
```{r}
plot.S1.5 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1324 & STOP_ID <= 1339) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,101) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.5
```
```{r}
plot.S1.6 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1339 & STOP_ID <= 1396) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,135) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.6
```
```{r}
plot.S1.7 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1396 & STOP_ID <= 1579) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,135) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.7
```
```{r}
plot.S1.8 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1579 & STOP_ID <= 1598) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,110) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.8
```
```{r}
plot.S1.9 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1598 & STOP_ID <= 1621) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,355) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.9
```
```{r}
plot.S1.10 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1621 & STOP_ID <= 1643) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,90) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.10
```
```{r}
plot.S1.11 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 1643 & STOP_ID <= 2024) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,60) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.11
```
```{r}
plot.S1.12 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 2024 & STOP_ID <= 2065) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,25) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.12
```
```{r}
plot.S1.13 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 2065 & STOP_ID <= 2296) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,60) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.13
```
```{r}
plot.S1.14 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 2296 & STOP_ID <= 2396) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.14
```
```{r}
plot.S1.15 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 2396 & STOP_ID <= 2456) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,50) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.15
```
```{r}
plot.S1.16 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 2456 & STOP_ID <= 2886) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.16
```
```{r}
plot.S1.17 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 2886 & STOP_ID <= 3296) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.17
```
```{r}
plot.S1.18 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 3296 & STOP_ID <= 3731) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.18
```
```{r}
plot.S1.19 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 3731 & STOP_ID <= 4192) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.19
```
```{r}
plot.S1.3931 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID == 3931) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,1500) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.3931
```
```{r}
plot.S1.20 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 4192 & STOP_ID <= 4613) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.20
```
```{r}
plot.S1.21 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 4613 & STOP_ID <= 4723) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,200) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.21
```
```{r}
plot.S1.22 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 4723 & STOP_ID <= 4933) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.22
```
```{r}
plot.S1.23 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 4933 & STOP_ID <= 5131) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,120) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.23
```
```{r}
plot.S1.24 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 5131 & STOP_ID <= 5404) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,75) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.24
```
```{r}
plot.S1.25 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 5404 & STOP_ID <= 5506) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,75) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.25
```
```{r}
plot.S1.26 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 5506 & STOP_ID <= 5617) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.26
```
```{r}
plot.S1.27 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 5617 & STOP_ID <= 5782) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,120) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.27
```
```{r}
plot.S1.28 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 5782 & STOP_ID <= 5915) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.28
```
```{r}
plot.S1.29 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 5915 & STOP_ID <= 6296) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,250) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.29
```
```{r}
plot.S1.30 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 6296 & STOP_ID <= 657) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,100) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.30
```
```{r}
plot.S1.31 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 657 & STOP_ID <= 673) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,50) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.31
```
```{r}
plot.S1.32 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 673 & STOP_ID <= 898) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,50) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.32
```
```{r}
plot.S1.33 <- 
  as.data.frame(data.y.18.S1) %>% 
  filter(STOP_ID > 898 & STOP_ID <= 913) %>%
  group_by(STOP_ID) %>% 
    ggplot(aes(x= MONTH_ID, y=AVERAGE_ON, group = 1)) + 
      geom_point() +
      plotTheme() +
      facet_wrap(~STOP_ID, scales="free", ncol=5) +
      ylim(0,50) +
      labs(title="Average Daily Ridership by Stop in Austin IDS1",
           subtitle="CapRemap Redesign Implemented Month in Blue", x="Month", y="Average Daily Ridership")+ 
      geom_vline(xintercept = 6, color = "blue")

plot.S1.33
```

Mapping ridership in May
```{r}
q5 <- function(variable) {as.factor(ntile(variable, 5))}

qBr <- function(df, variable, rnd) {
  if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                          c(.01,.2,.4,.6,.8), na.rm=T))
  } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]]), digits = 3),
                 c(.01,.2,.4,.6,.8), na.rm=T)
  }
}

palette5 <- c("#FA7800","#C48C04","#8FA108","#5AB60C","#25CB10")

mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}
data.y.18.5 <- data.y18 %>%
  subset(data.y18$MONTH_ID == 5)

data.y.18.5.sf <-
  data.y.18.5 %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, agr = "constant") %>%
  st_transform(2278)

ggplot()+
  geom_sf(data=TrusteeDist,fill="grey20")+
  geom_sf(data=data.y.18.5.sf,aes(colour=q5(AVERAGE_ON)),show.legend = "point",size = .75)+
  scale_color_manual(values=palette5,
                     labels=qBr(data.y.18.5.sf,"AVERAGE_ON"),
                     name="Quintile\nBreaks")+
  labs(title="Ridership in May, Austin")+
  mapTheme()
```
Mapping ridership in June
```{r}
data.y.18.6 <- data.y18 %>%
  subset(data.y18$MONTH_ID == 6)

data.y.18.6.sf <-
  data.y.18.6 %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326, agr = "constant") %>%
  st_transform(2278)

ggplot()+
  geom_sf(data=TrusteeDist,fill="grey20")+
  geom_sf(data=data.y.18.6.sf,aes(colour=q5(AVERAGE_ON)),show.legend = "point",size = .75)+
  scale_color_manual(values=palette5,
                     labels=qBr(data.y.18.6.sf,"AVERAGE_ON"),
                     name="Quintile\nBreaks")+
  labs(title="Ridership in June, Austin")+
  mapTheme()
```

